---
title: "Data Analysis of Life Sciences Grants via GRIST API"
output: 
  html_notebook:
    code_folding: hide
date: 2020-09-21
---
This project aims to source data about awarded grants in the life sciences and to conduct analysis on this data.

Data Source: Grants RESTFul (GRIST) API from Europe PMC
This provides coverage of funding from BBSRC UK; CRUK; NIHR; ERC; MRC UK; Wellcome Trust; WHO and more.

The ideal workflow will be like so:

1. Input search term
2. Python3's urllib will utilise GRIST API to spit out .json file for each 25 results from search term
3. SQLite3 database will be used to cache each .json file
4. Once all .json files have been cached, Python3 will run through each .json file & strip out details from each grant entry:
  a) Full Name
  b) Institution
  c) Start Date
  d) End Date
  e) Grant Title
  f) Grant Abstract
  h) Grant Source
  i) Grant Type
  j) Grant Worth + Currency
  
and store them in a Numpy array with each row reflecting separate ID i.e grant entry.

5. Once, Numpy array has been constructed as so, it will be converted to a csv named with the particular 'search term' used. which can then be read into a tibble in R.

6. Analysis of that particular 'search term' data can then be carried out to answer the following questions:
- Who is the most well-funded for that 'search term' & when?
- Which research institution is most well-funded for that 'search term' and when?
- How much money has that 'search term' been funded over the past as a running total?


7. Comparative analysis can also be conducted between tibbles for 'search terms'
- How much money per grant is awarded between the different 'search terms'?
- How much total money is awarded over the years for different 'search terms'?
- How many grants have been awarded over the years for different 'search terms'?

8. Data will be outputted in the form of a flexdashboard::flex_dashboard or Shiny web app.


```{r setup, include = FALSE}
library(tidyverse)
library(plotly)
library(quantmod)
library(lubridate)

knitr::opts_chunk$set(fig.width = 12, fig.asp = 0.618, fig.align = "center") 
```

Note that before you run the following code for data analysis, you should have run the GRIST.py script to output .tsv files of interest to the folder ./output_tsv

You also, have to physically run all the code using 'Run' button first and then output via Knitr.

The below code will import the .tsv files found within the ./output_tsv folder as individual tibbles.

```{r import}
filenames <- list.files("./output_tsv")

for (i in seq_along(filenames)){
  file <- str_extract(filenames[i],".+\\.") %>%
    str_extract(".+[^\\.]")
  
  tsv <- read_tsv(str_c("./output_tsv/",filenames[i]), 
                     col_names = c("ID","Name","Institution","Department", "Start Date","End Date","Title","Abstract", "Funder","Grant Type","Amount", "Currency"), 
                     col_types = "icccDDccccdc",
                     na="None",
                     quote="")
  
  intermediate <- file
  assign(intermediate,tsv)
  rm(tsv,file,intermediate) 
}
```


The below code will do some accessory work to compile all tibbles to iterate over all the tibbles & convert all currency values to GBP as per FX rates of `r Sys.time()`.

The tsv files also seem to have duplicate rows for some so this will have to be resolved.

```{r acc, message=FALSE, warning=FALSE}
search_terms <- vector("character",length(filenames))
for (i in seq_along(filenames)){
  search_terms[i] <- str_extract(filenames[i],".+\\.") %>%
    str_extract(".+[^\\.]")
}

currency_denoter <- function(amount_input,currency_input) { #Converts the currency for single row
  getFX(str_c(currency_input,"/GBP"),Sys.Date()-1)
  varname <- str_c(currency_input,"GBP")
  rate <- eval(parse(text=varname))
  gbp_out <- as.double(amount_input * rate[1,1])
  return(gbp_out)
}

currency_standardiser <- function(input_tb){
  standardised_rows <- input_tb %>%
    distinct(Abstract, .keep_all=TRUE) %>% #This is to remove duplicate rows
    filter(Currency != "GBP" & !is.na(Currency)) %>%
    mutate(Amount = map2_dbl(Amount,Currency,currency_denoter)) %>%
    mutate(Currency = "GBP")
  
  standardised_tb_V2 <- input_tb %>%
    distinct(Abstract, .keep_all=TRUE) %>% #This is to remove duplicate rows
    filter(Currency == "GBP" | is.na(Currency)) %>%
    full_join(standardised_rows) %>%
    arrange(ID)
}

for (i in seq_along(search_terms)){ #This code is horribly inefficient but it works so...
  relevant_tb <- eval(parse(text=str_c("`",search_terms[i],"`")))
  standardised_tb <- currency_standardiser(relevant_tb) %>%
    mutate(Search_Term = search_terms[i]) #Adds an extra column denoting the particular search term
  assign(search_terms[i],standardised_tb) # This assigns the new, standardised tibble to the same variable name as the                                             previous one.
  rm(standardised_tb,relevant_tb)
}

rm(filenames)
```

The below code will answer the question of **who are the x most well-funded PI ?** (with arguments for top x as well as end date to filter) and **for the top x most well-funded PIs, what was the period of the grant & how much was each worth ?**.

However, note that using end date to filter may remove entries which have missing end dates and end date cutoff must be inputted as a string.

It will answer this question for each of the search terms, outputting for each search term

1. A sorted bar chart of x PIs
2. A line graph of total grant money active at any one time for each of x most well-funded PIs.

The plots are themselves interactive:
- Hover over each point or bar to gain quantitative information
- Use the toolbar functions to zoom in/out
- Click on legend entries to remove certain data points from the graph.


```{r PI_Analysis}

top_x_PIs <- function(input_tb,x_to_keep=length(input_tb), end_date_cutoff = as.character(min(input_tb$`End Date`, na.rm=TRUE))){ # End date used as cutoff instead of start date because more results have NA as start date cf. end date.
  top_PIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Name) %>% 
    mutate(Total_Amount = sum(Amount)) %>%
    select(2,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct(Name, .keep_all= TRUE) %>% # This is to remove duplicate entries for certain PIs
    ungroup()
  
  top_x <- slice_head(top_PIs, n=x_to_keep) %>%
    mutate(Ordered_Name = reorder(Name,Total_Amount))
  #biggest_moneymaker <- str(top_x[[1]][[1]])
  #biggest_earnings <- str(top_x[[1]][[2]])
  #tb_name <- deparse(substitute(input_tb)) # code to grab the name of the variable itself
  
  
  plot <- ggplot(top_x,aes(Total_Amount, Ordered_Name))+
    geom_col() +
    labs(x="Total GBP Awarded",y="Name of PI", 
         title=str_c("Top",as.character(x_to_keep),"Well-Funded PIs for", str_to_title(top_x$Search_Term[[1]]), sep=" "), caption="Europe PMC GRIST API", subtitle = str_c("n=",as.character(nrow(input_tb))))
  
  plot <- ggplotly(plot, tooltip=c("Total_Amount")) %>%
    layout(title = list(text = paste0(str_c("Top",as.character(x_to_keep),"Well-Funded PIs for", str_to_title(top_x$Search_Term[[1]]), sep=" "),
                                    '<br>',
                                    '<sup>',
                                    str_c("n=",as.character(nrow(input_tb))),
                                    '</sup>')))
  
  print(plot)
  
  return()
}



total_running_money_top_x_PIs <- function(input_tb, x_to_keep = length(input_tb),
                                          end_date_cutoff = as.character(min(input_tb$`End Date`, na.rm=TRUE))){
  
   top_x_PIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Name) %>%
    mutate(Total_Amount = sum(Amount)) %>%
    select(2,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct(Name, .keep_all= TRUE) %>%
    ungroup()
  
  top_x <- slice_head(top_x_PIs, n=x_to_keep)
  
  
  top_running_money_PIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Name) %>%
    mutate(Total_Amount = sum(Amount)) %>%
    select(2,5:6,11,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct() %>%
    ungroup()

  top_x_amount <- semi_join(top_running_money_PIs, top_x, by="Name") %>%
    group_by(Name) %>%
    pivot_longer(c(`Start Date`,`End Date`), names_to = "Dates", values_to = "Values")
  
  
  total_at_any_time <- vector("numeric",nrow(top_x_amount))
  
  for (i in seq(1,nrow(top_x_amount))){
    
    if(i == nrow(top_x_amount) || i == 1) {
      total_at_any_time[[i]] <- top_x_amount$Amount[[i]]
    }
    
    else{
      
      if(top_x_amount$Dates[[i]] == "Start Date" && top_x_amount$Dates[[i+1]] == "End Date" && top_x_amount$Dates[[i-1]] == "End Date"){
        total_at_any_time[[i]] <- top_x_amount$Amount[[i]]
      }
      
      if(top_x_amount$Dates[[i]] == "End Date" && top_x_amount$Dates[[i+1]] == "Start Date" && top_x_amount$Dates[[i-1]] == "Start Date"){
        total_at_any_time[[i]] <- top_x_amount$Amount[[i]]
      }
      
      # i.e the above two conditions are normal i.e start, end, start, end ... for the Dates column
      
      if(top_x_amount$Dates[[i]] == "Start Date" && top_x_amount$Dates[[i+1]] == "Start Date" && top_x_amount$Dates[[i-1]] != "Start Date") {
        total_at_any_time[[i]] <- top_x_amount$Amount[[i]]
        total_at_any_time[[i+1]] <- total_at_any_time[[i]] + top_x_amount$Amount[[i+1]]
      }
      
      if(top_x_amount$Dates[[i]] == "Start Date" && top_x_amount$Dates[[i+1]] == "Start Date" && top_x_amount$Dates[[i-1]] == "Start Date") {
        total_at_any_time[[i+1]] <- total_at_any_time[[i]] + top_x_amount$Amount[[i+1]]
      }
      
      if(top_x_amount$Dates[[i]] == "Start Date" && top_x_amount$Dates[[i+1]] == "End Date" && top_x_amount$Dates[[i-1]] == "Start Date"){
        total_at_any_time[[i+1]] <- total_at_any_time[[i]] - top_x_amount$Amount[[i+1]] # Junction/peak point for start
      }
      
      if(top_x_amount$Dates[[i]] == "End Date" && top_x_amount$Dates[[i+1]] == "End Date" && top_x_amount$Dates[[i-1]] == "Start Date") {
        total_at_any_time[[i+1]] <- total_at_any_time[[i]] - top_x_amount$Amount[[i+1]] #Junction/peak point for end
      }
      
      if(top_x_amount$Dates[[i]] == "End Date" && top_x_amount$Dates[[i+1]] == "End Date" && top_x_amount$Dates[[i-1]] == "End Date") {
        total_at_any_time[[i+1]] <- total_at_any_time[[i]] - top_x_amount$Amount[[i+1]] #Only applies for >= 3 grants active
      }
      
      
      
    } # Note that the above statements write the value of the one in front.
    
  }
  
  top_x_amount <- cbind(top_x_amount, Rolling_Amount = total_at_any_time)
  
  plot <- ggplot(top_x_amount)+
      scale_x_date(name = "Date", date_breaks="2 year", date_labels="%Y")+
      geom_point(aes(x=Values,y=Rolling_Amount, col=Name))+
      geom_step(aes(x=Values, y=Rolling_Amount, col=Name))+
      labs(y="Total Running Amount of GBP Awarded",x="Year", 
         title=str_c("Running Grant GBP for Top",as.character(x_to_keep),"Well-Funded PIs for", str_to_title(top_x_amount$Search_Term[[1]]), sep=" "), caption="Europe PMC GRIST API",subtitle = str_c("n=",as.character(nrow(input_tb))))
    
  plot <- ggplotly(plot) %>%
    layout(title = list(text = paste0(str_c("Running Grant GBP for Top",as.character(x_to_keep),"Well-Funded PIs for", str_to_title(top_x_amount$Search_Term[[1]]), sep=" "),
                                    '<br>',
                                    '<sup>',
                                    str_c("n=",as.character(nrow(input_tb))),
                                    '</sup>')))
  print(plot)
  
  return()
}

for (i in seq_along(search_terms)){
  relevant_tb <- eval(parse(text=str_c("`",search_terms[i],"`")))
  sorted_bar_chart <- top_x_PIs(relevant_tb,20,"2010-01-01") #default arguments = 20 PIs and for finish date > 2010
}

for (i in seq_along(search_terms)){
  relevant_tb <- eval(parse(text=str_c("`",search_terms[i],"`")))
  sorted_bar_chart <- total_running_money_top_x_PIs(relevant_tb,20,"2010-01-01") #default arguments = 20 PIs and for finish date > 2010
}
```

The below code will answer the question of **who are the x most well-funded research institutions ** (with arguments for top x as well as end date to filter) and **for the top x most well-funded research institutions, what was the period of the grant & how much was each worth ?**.

However, note that using end date to filter may remove entries which have missing end dates and end date cutoff must be inputted as a string.

It will answer this question for each of the search terms, outputting for each search term

1. A sorted bar chart of x research institutions.
2. A line graph of total grant money active at any one time for each of x most well-funded research institutions.

```{r RI_Analysis}

top_x_RIs <- function(input_tb,x_to_keep=length(input_tb), end_date_cutoff = as.character(min(input_tb$`End Date`, na.rm=TRUE))){ # End date used as cutoff instead of start date because more results have NA as start date cf. end date.
  top_x_RIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Institution) %>% 
    mutate(Total_Amount = sum(Amount)) %>%
    select(Institution,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct(Institution, .keep_all= TRUE) %>%
    ungroup()
  
  top_x <- slice_head(top_x_RIs, n=x_to_keep)
  
  plot <- ggplot(top_x,aes(Total_Amount, reorder(Institution, Total_Amount)))+
    geom_col() +
    labs(x="Total GBP Awarded",y="Name of Research Institution", 
         title=str_c("Top",as.character(x_to_keep),"Well-Funded RIs for", str_to_title(top_x$Search_Term[[1]]), sep=" "), caption="Europe PMC GRIST API", subtitle = str_c("n=",as.character(nrow(input_tb))))
  
  plot <- ggplotly(plot,tooltip=c("Total_Amount"))
  print(plot)
  
  
  return()
}


total_running_money_top_x_RIs <- function(input_tb, x_to_keep = length(input_tb),
                                          end_date_cutoff = as.character(min(input_tb$`End Date`, na.rm=TRUE))){
  
   top_x_RIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Institution) %>% 
    mutate(Total_Amount = sum(Amount)) %>%
    select(Name,Institution,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct(Institution, .keep_all= TRUE) %>%
    ungroup()
  
  top_x <- slice_head(top_x_RIs, n=x_to_keep)
  
  
  top_running_money_RIs <- input_tb %>%
    filter(`End Date` >= ymd(end_date_cutoff)) %>%
    group_by(Institution) %>%
    mutate(Total_Amount = sum(Amount)) %>%
    select(Name,Institution,5:6,11,Total_Amount,Search_Term) %>%
    arrange(desc(Total_Amount)) %>%
    distinct() %>%
    ungroup()

  top_x_amount <- semi_join(top_running_money_RIs, top_x, by="Institution") %>%
    group_by(Institution) %>%
    pivot_longer(c(`Start Date`,`End Date`), names_to = "Dates", values_to = "Values")

  
  plot <- ggplot(top_x_amount)+
      scale_x_date(name = "Date", date_breaks="2 year", date_labels="%Y")+
      geom_point(aes(x=Values,y=Amount, col=Institution))+
      geom_step(aes(x=Values, y=Amount, col=Institution, group=Name))+ #It will show each grant as a line from start to end.
      labs(y="Grant Award Amount /GBP",x="Year", 
         title=str_c("Grant Amounts for Top",as.character(x_to_keep),"Well-Funded RIs for", str_to_title(top_x_amount$Search_Term[[1]]), sep=" "), caption="Europe PMC GRIST API", subtitle = str_c("n=",as.character(nrow(input_tb))))
     # ggrepel::geom_text_repel(aes(x=Values, y=Amount,label=Name),data=top_x_amount)
  
  plot <- ggplotly(plot)
  print(plot)
    
  
  return()
}

for (i in seq_along(search_terms)){
  relevant_tb <- eval(parse(text=str_c("`",search_terms[i],"`")))
  sorted_bar_chart <- top_x_RIs(relevant_tb,10,"2010-01-01") #default arguments = 5 RIs and for entire date range
}

for (i in seq_along(search_terms)){
  relevant_tb <- eval(parse(text=str_c("`",search_terms[i],"`")))
  sorted_bar_chart <- total_running_money_top_x_RIs(relevant_tb,5,"2010-01-01") #default arguments = 5 RIs and for entire date range
}

```

The below code will answer the question of **how much total money has that 'search term' been funded over the past?**

It will answer this question for each of the search terms, outputting for each search term

1. A line graph showing the cumulative sum of money over time
2. A bar graph showing the total money granted in the past 2, 5, 10 years

